---
title: "Patients age 50-80 with 20+ pack years"
author: "Jamie Medina"
date: "2/9/2021"
output: html_document
---


```{r background, echo=FALSE, include=FALSE}
###################################
# Task: Predict 
# Created: 012221
# Due date: Monday 25, 2021 or Tuesday 26, 2021
#####################################
library(precrec)
library(tidyverse)
library(caret)
library(pROC)
library(readxl)
library(DT)
library(recipes)
library(devtools)
library(data.table)
library(here)
```

```{r setup, include=FALSE}
#Clean out list section
rm(list=ls());gc()
```


## Reviewer #2 request
 - "we are looking at reviewer #2 and he is asking us to re-train the model with patients age 50-80 and smoking history of 20 packyears and above. If I provide you with the list of these patients could you retrain the standard model that we have been running for LUCAS (ratios, PCs of z score)? Thank you!" - Slack

# Patients age 50-80, smoking history of 20 packyears and above
- 203 individuals
- 91 Cancers, 112 Healthy

# the below analysis also contains updated filtered Patients age 50-80, smoking history of 20+ pys

# Content also includes report4.rmd re-run of model_seq models for ROC plotting in report8.rmd

# Further filtering was done by Dimitrios to exclude prior cancers - slack message - "@Jmedina thank you for running this last iteration of LUCAS samples? These are pts of age 50-80, packyears>20, excluding pts with prior cancer and including patients with no baseline cancer and future cancers as healthies" - Metadata indicated 157 samples in this analysis.

```{r model_seq, echo = FALSE, message=FALSE, warning=FALSE}
#load meta and fragment data
lucas.updated <- read_csv(here("data","lucas_5mbs_delfi473.csv"))
lucas.metadata <- read_excel(here("data","LUCAS_pts_50_80_py20_excluding prior cancers.xlsx"))

lucas <- inner_join(lucas.updated, lucas.metadata, by = c("id"="id"))
lucas <- lucas %>% filter(chr != "chrX")

length(unique(lucas$id))
#compile ratio and coverage
lucas <- lucas %>% group_by(id) %>%
  mutate(ratio.cor = short.cor/ long.cor,
         ratio.scaled = scale(ratio.cor),
         ratiovar = factor(paste0("ratio_", bin), paste0("ratio_",1:473)),
         cov.cor = short.cor+long.cor,
         cov.scaled = scale(cov.cor),
         covvar = factor(paste0("cov_", bin), paste0("cov_",1:473)),
         status=ifelse(type=="Healthy", "Healthy", "Cancer"))

#zscores
lucaz <- read_csv(here("data","training-set.csv"))
lucaz <- lucaz %>% select(id, contains("zscore"))
colnames(lucaz)
lucaz

###########################################
# Append Nature Features
###########################################

features.l <- lucas %>% ungroup() %>%
  select(Patient, id, type, status, bin, ratio.scaled, ratiovar, covvar, short.cor, cov.scaled, contains("Z Score"))

#pivot features
features <- features.l %>% unite("id", id, type) %>% pivot_wider(id, names_from=ratiovar, values_from=c(ratio.scaled)) %>% rowwise() %>% mutate(type = strsplit(id, "_")[[1]][2], id2 = strsplit(id,"_")[[1]][1]) %>% select(id2, type, everything(), -id)

#append in zscores
features <- cbind(features, as.matrix(lucaz %>% filter(id %in% features$id2) %>%
                                               group_by() %>%
                                               select(contains("zscore")) %>%
                                               distinct()))

dim(features)


#######
# Model implementing newer guidelines
# Need to change zscores to
ratio_seq <- recipe(type ~ ., data= features) %>%
  update_role(id2, new_role = "ID") %>%
  step_pca(starts_with("ratio_"), prefix = "ratio_pc_", threshold = 0.90) %>%
  step_corr(all_predictors(), threshold = 0.95) %>%
  step_nzv(all_predictors())

glmnetGrid <- expand.grid(alpha = 1, lambda = 10^seq(-5, -1, length.out = 100))

### Model Training
set.seed(1234)
ctrl <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 10,
                     verboseIter = TRUE,
                     savePredictions="final",
                     classProbs=TRUE,
                     index = createMultiFolds(features$type, 5, 10),
                     summaryFunction = twoClassSummary)

model_seq <- caret::train(ratio_seq,
                               data = features,
                                method = "glmnet",
                               tuneGrid = glmnetGrid,
                               trControl = ctrl)

saveRDS(model_seq, "../models_c3/model_seq_glm.rds")


###predictions

pred.cov <- model_seq$pred %>%
    filter(alpha==model_seq$bestTune$alpha, lambda==model_seq$bestTune$lambda) %>%
    group_by(rowIndex) %>% dplyr::summarize(obs=obs[1], score.seq=mean(Cancer))

pred.cov %>% filter(obs=="Healthy") %>% arrange(desc(score.seq))
cutoff99.model_opt <- (pred.cov %>% filter(obs=="Healthy") %>% arrange(desc(score.seq)))$score.seq[2]
cutoff95.model_opt <- (pred.cov%>% filter(obs=="Healthy") %>% arrange(desc(score.seq)))$score.seq[5]
cutoff90.model_opt <- (pred.cov %>% filter(obs=="Healthy") %>% arrange(desc(score.seq)))$score.seq[9]
cutoff80.model_opt <- (pred.cov %>% filter(obs=="Healthy") %>% arrange(desc(score.seq)))$score.seq[18]
pred.cov.final <- pred.cov %>% mutate(detected99 =
                                                              ifelse(score.seq > cutoff99.model_opt, "Cancer", "Healthy"),
                                                            detected95 =
                                                              ifelse(score.seq> cutoff95.model_opt, "Cancer", "Healthy"),
                                                            detected90 =
                                                              ifelse(score.seq > cutoff90.model_opt, "Cancer", "Healthy"),
                                                            detected80 =
                                                              ifelse(score.seq> cutoff80.model_opt, "Cancer", "Healthy"))

pred.cov.final$id <- (features %>% select(id2) %>% distinct())$id2
final <- inner_join(pred.cov.final, lucas.metadata, by =c("id"="id"))
final$type <- NULL
names(final)
write_csv(final, "prediction_lucas_c3.csv")





```


