---
title: "ML models evaluated"
author: "Rob Scharpf"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(here)
```


```{r models}
here("data", "models.csv") %>%
    read_csv() %>%
    kbl() %>%
    kable_styling(full_width=FALSE)
```

I believe the only two models of interest are `seq` and `full_lasso2`.  For the `seq` model, we had three different training/testing cohorts:

1. the LUCAS cohort (158 non-cancer, 129 cancer)
2. the LUCAS cohort excluding prior cancers
3. the LUCAS cohort excluding prior cancers and limited to 50-80 yo smokers.


We should define a `cohort1`, `cohort2`, and `cohort3` indicator variables in the summarized experiment object in the `rlucas` package. These variables should take the value 1 for samples included in the cohort and 0 otherwise.  The `seq` model based on cohort 1 above was externally validated.  We should include the code for the validation in the workflowr project as well.
