---
title: "GC model"
site: workflowr::wflow_site
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: console
---

```{r caching, echo=FALSE}
knitr::opts_chunk$set(autodep = TRUE, warning=FALSE)
```


```{r packages, message=FALSE, warning=FALSE}
library(RColorBrewer)
library(SummarizedExperiment)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(here)
```


```{r gc_models}
odir <- here("data", "gc_correction.R")
loess_model <- here(odir, "loess_model.rds") %>%
    readRDS()
frag_model <- here(odir, "frag_model.rds") %>%
    readRDS()
```


```{r loess_ratios}
loess_ratios <- loess_model %>%
    ungroup() %>%
    select(-c(gc, count)) %>%
    pivot_wider(id_cols=c(lab_id, bin),
                names_from=size, values_from=loess) %>%
    mutate(ratio=short/long) %>%
    mutate(method="Bin-level\nGC model")
```

```{r frag_ratios}
frag_model %>%
    dplyr::rename(lab_id=id) %>%
    group_by(lab_id) %>%
    summarize(nbin=length(unique(bin)),
              .groups="drop")

frag_ratios <- frag_model %>%
    dplyr::rename(lab_id=id) %>%
    select(lab_id, bin, short, long, method) %>%
    mutate(ratio=short/long) %>%
    mutate(method="Fragment-level\nGC model")

mlevels <- c("Bin-level\nGC model",
             "Fragment-level\nGC model")
combined <- bind_rows(loess_ratios, frag_ratios) %>%
    mutate(method=factor(method, mlevels))
```

Attach metadata on the samples and metadata on the rows.

```{r bin_metadata}
##data(pcr_experiment, package="pcr.data")
load(here("data", "gc_correction.R", "pcr_experiment.rda"))
rr <- rowRanges(pcr_experiment) %>%
    as_tibble() %>%
    select(bin, arm)
combined2 <- left_join(combined, rr, by="bin")
```

```{r sample_metadata}
coldat <- colData(pcr_experiment) %>%
    as_tibble() %>%
    dplyr::rename(lab_id=sample) %>%
    select(lab_id, pcr_cycles)
combined3 <- left_join(combined2, coldat, by="lab_id")
```

# Fragmentation profiles

```{r profiles, fig.width=12, fig.height=8, dev=c("png", "pdf")}
zero_cycle <- combined3 %>%
    filter(pcr_cycles==0, method==levels(method)[2]) %>%
    ##filter(arm %in% levels(arm)[1:5]) %>%
    mutate(method="Zero-cycle\nNo GC correction")
combined4 <- combined3 %>%
    mutate(method=as.character(method)) %>%
    bind_rows(zero_cycle) %>%
    mutate(method=factor(method, c("Zero-cycle\nNo GC correction", mlevels)),
           pcr_cycles=as.character(pcr_cycles),
           pcr_cycles=factor(pcr_cycles, c("0", "4", "12")))

pal <- brewer.pal(n=3, "Dark2") %>%
    setNames(c("0", "4", "12"))
pal[1] <- "black"

combined4 %>%
    filter(arm %in% levels(arm)[1:6]) %>%
    ggplot(aes(bin, ratio, group=lab_id)) +
    geom_line(aes(color=pcr_cycles)) +
    facet_grid(method~arm, space="free", scales="free_x") +
    scale_color_manual(values=pal) +
    theme_bw(base_size=17) +
    theme(panel.grid=element_blank(),
          strip.text.y=element_text(hjust=0, angle=0),
          strip.background=element_blank(),
          legend.position="bottom") +
    guides(color=guide_legend(title="PCR cycles")) +
    ylab("Fragmentation profile")
```

Fragmentation profiles and models for GC correction. Row 1: Short to long ratios for a zero-cycle PCR experiment without GC correction.  As PCR can preferentially amplify fragments by length and GC content, we used a loess model to correct the number of short fragments and number of long fragments in 5 Mb bins along the genome for the 0, 4, and 12-cycle samples.   While the bin-level model reduces biases in coverage from length and GC content by PCR (Row 2), the fragmentation profiles no longer resemble the zero-cycle fragmentation profile. Row 3:  A non-parametric fragment-level model that normalizes coverage by fragment length and GC content to the zero-cycle PCR target distribution.
